---
title: "Data 606 - HW3"
author: "Baron Curtin"
date: "March 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(magrittr)
library(DATA606)
```



## 3.2 I. What percent of a standard normal distribution N(µ =0,= 1) is found in each region? Be sure to draw a graph.

```{r 3.2a}
m <- 0
sd <- 1
z <- -1.13

# find x
x <- (z * sd * m) %>%
  print

# probability x > -1.13
1 - pnorm(x, mean = 0, sd = 1)

# probability plot
normalPlot(mean = 0, sd = 1, bounds = c(x, 1e+06), tails = FALSE)
```

```{r 3.2b}
m <- 0
sd <- 1
z <- 0.18

# finding x
x <- (z * sd * m) %>%
  print

# probability x < .18
pnorm(x, mean = 0, sd = 1)

# probability plot
normalPlot(mean = 0, sd = 1, bounds = c(-1e+06, x), tails = FALSE)
```

```{r 3.2c}
m <- 0
sd <- 1
z <- 8

# finding x
x <- (z * sd * m) %>%
  print

# probability x > 8
1 - pnorm(x, mean = 0, sd = 1)

# probability plot
normalPlot(mean = 0, sd = 1, bounds = c(x, 1e+06), tails = FALSE)
```

```{r 3.2d}
m <- 0
sd <- 1
z <- .5

# finding x
x <- (z * sd * m) %>%
  print

# probability |x| < .5
x1 <- pnorm(-x, mean = 0, sd = 1)
x2 <- pnorm(x, mean = 0, sd = 1)
print(x2 - x1)

# probability plot
normalPlot(mean = 0, sd = 1, bounds = c(-x, x), tails = FALSE)
```

## 3.4
  * a.) Group Men, Ages 30 - 34: N(4313, 583)
        Group Women, Ages 25-29: N(5261, 807)
  * b.)
```{r 3.4b}
# leo zscore
ltime <- 4948
lm <- 4313
lsd <- 583

zleo <- ((ltime - lm)/lsd) %>%
  print

# mary zscore
mtime <- 5261
mm <- 4313
msd <- 807

zmary <- ((mtime - mm)/msd) %>%
  print
```

  * c.) Mary performed better in her group than Leo because she was in the top 12% of her group as opposed to Leo at 13.80% in his.
```{r 3.4c}
# leo rank
leo <- 1 - pnorm(ltime, mean = lm, sd = lsd) %>%
  print

# mary rank
mary <- 1 - pnorm(mtime, mean = mm, sd = msd) %>%
  print
```

  * d.)
```{r 3.4d}
pnorm(ltime, mean = lm, sd = lsd) %>%
  print
```

  * e.)
```{r}
pnorm(mtime, mean = mm, sd = msd) %>%
  print
```

  * f.) The answer to B would not change because Z-scores can be calculated for non-normal distributions. Part E would be difficult to calculate do to it not being a normal distribution.

## 3.18 The data does follow the 68-95-99.7 rule for the most part.
```{r 3.18}
heights <- c(54,55,56,56,57,58,58,59,60,60,60,61,61,62,62,63,63,63,64,65,65,67,67,69,73)
m <- 61.52
sd <- 4.58

zScores <- (heights - m) / sd
hist(zScores, col="white", freq=F, xlim=c(-5, 5))
curve(dnorm, -3, 3, add=T, col="blue") # 99.7% Rule
curve(dnorm,-2,2,add=T, col="red") # 95% Rule
curve(dnorm,-1,1,add=T, col="orange") # 68% Rules
```

## 3.22
  * a.) The probability is almost 0% that the 10th transitor will be produced with a defect.
```{r 3.22a}
# Rate of success and failure definition
pf <- 0.02
ps <- 1 - pf
n <- 10

# This is a geometric distribution
round(dgeom(n, ps), 4)

round(ps * (1 - ps)^(n - 1), 4) %>%
  print
```

  * b.) There is a 13.26% chance the machines produces zero defective transitors.
```{r 3.22b}
# Rate of success and failure definition
pf <- 0.02
ps <- 1 - pf
n <- 100

# This is a geometric distribution
round(ps^n, 4) %>%
  print
```

  * c.) On average, there should be 50 transitors produced before the first defect, with a standard deviation of 49.50.
```{r 3.22c}
# Expected value of a geometric distribution
pf <- 0.02
Ex <- (1/pf) %>%
  print

# standard deviation
sd <- (((1 - pf)/pf^2)^(1/2)) %>%
  print
```

  * d.) On average, there should be 20 transitors before the first defect, with a standard deviation of 19.50.
```{r 3.22d}
# Expected value of a geometric distribution
pf <- 0.05
Ex <- (1/pf) %>%
  print

# standard deviation
sd <- (((1 - pf)/pf^2)^(1/2)) %>%
  print
```

  * e.) Increasing the probability of failure decreases the trials before a defect and therefore the waiting time before a failure (success) and standard deviation are smaller.

## 3.38
  * a.)
```{r 3.38a}
n <- 3
k <- 2
pboy <- 0.51
pboy2 <- (choose(n, k) * (1 - pboy)^(n - k) * (pboy)^k) %>%
  print
```

  * b.)
```{r 3.38b}
children <- data.frame(c("BBG", "BGB", "GBB"))
children$p <- c(pboy * pboy * (1 - pboy), pboy * (1 - pboy) * pboy, (1 - pboy) * 
    pboy * pboy)
names(children) <- c("Kids", "p")

sump <- sum(children$p) %>%
  print

(pboy2 - sump) %>%
  print

children %>%
  print
```

  * c.) The second method will be more tedious because a combination of 56 different possibilities will have to be made. This would create a great deal of redundancy.

## 3.42
  * a.)
```{r 3.42a}
# Negative Binomial distribution
p <- 0.15
n <- 10
k <- 3

(choose(n - 1, k - 1) * (1 - p)^(n - k) * p^k) %>%
  print
```

  * b.) The probability that her 10th serve will be successful is 15% since they are independent events.
  * c.) The probability are different because part A was the calculation of a specific scenario with 3 successes in 10 attempts. The probability of B is calculated for only one attempt.



