---
title: "Data 606 - Project"
author: "Baron Curtin"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    df_print: paged
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(collapse = TRUE)
pkgs <- c("tidyverse", "magrittr", "kableExtra", "knitr")
lapply(pkgs, require, character.only=TRUE)
```

### Part 1 - Introduction

There is growing sentiment within society that instances of hate are increasing. In the 10 days following the 2016 election, nearly 900 hate incidents were reported to the Southern Poverty Law Center (Majumder 2017). Thats an average of 90 hate crimes per day. For contrast, 36,000 hate crimes were reported to the FBI from 2010-2015, or 16 hate crimes per day on average.

The data we will be using for the analysis is a compilation of data from the Kaiser Family Foundation, the US Census Bureau, the US Election Project, Southern Poverty Law Center, and the FBI. The data is posted on FiveThirtyEight's GitHub in a CSV format. There are some caveats associated with the data: 1) the federal government doesn't track hate crimes systematically and thus some of the data is reliant upon media accounts and people's self-reports, 2) the FBI collects hate crime data from law enforcement agencies but the data is submitted voluntarily, 3) the FBI uses the Uniform Crime Reporting Program (UCR) to collect crime data, but the UCR only collects data on *prosecutable* hate crimes, 4) heightened news coverage of hate incidents after the election may have caused people to report hate incidents more frequently due to awareness bias (Majumder 2017). All of these things contribute to bias baked in to the data being used for analysis.

The question we will be looking at in this analysis is:  

> Are there more annual hate crimes per 100,00 population in areas where the greater share of population voted for Trump in 2016?


### Part 2 - Data

```{r dataload, results='hide', message=FALSE}
hateCrimes <- read_csv("https://raw.githubusercontent.com/baroncurtin2/data606/master/project/data/hate_crimes.csv")
regionMap <- read_csv("https://raw.githubusercontent.com/baroncurtin2/data606/master/project/data/region_mapping.csv") %>%
  rename_all(funs(str_to_lower(.)))
```

Some manipulations will need to be done to the data. *hateCrimes* will be joined to *regionMap* to add the columns: Region, Division, and State.Code. The following code will join the two datasets and create an additional categorical variable to denote Trump support

```{r}
data <- hateCrimes %>%
  # join two datasets
  left_join(regionMap, by = "state") %>%
  # add trump_support variables
  mutate(support = case_when(
    share_voters_voted_trump > .5 ~ 'Trump',
    share_voters_voted_trump < .5 ~ 'Clinton',
    TRUE ~ 'Split'
  ))
```

The decision here was made to create a column titled *support*. The condition was based on the percentage share of voters. A share of voters greater than 50% received a value of 'Trump', less than 50% received 'Clinton', and 50% received 'Split'.

We can further trim some of the "fat" in the dataset to only keep the columns we are analyzing. We will keep the median household income as a comparison for 'Trump Support'. The original FiveThirtyEight article quoted income being the greatest predictor of hate crimes. We can also rename some columns for easier reference later
```{r}
data %<>%
  select(-(3:9)) %>%
  # rename column
  rename(median_income = median_household_income,
         trump_share = share_voters_voted_trump,
         hate_crime_proportion = hate_crimes_per_100k_splc,
         avg_hatecrimes = avg_hatecrimes_per_100k_fbi)
```


### Part 3 - Exploratory Data Analysis

Lets get some general information on the dataset
```{r}
str(data)
```
We are able to see that there are 51 observations, and only 9 variables in the trimmed data set. We may want to convert some of the columns to factors in later analysis.


#### Distribution

```{r}
plotSetup <- ggplot(data)

plotSetup +
  geom_bar(mapping = aes(x = support, 
                         color = support, 
                         fill = support))
```
Nationwide, there Trump appears to be the less popular option

We are get more granular by looking at the distribution by region
```{r}
plotSetup +
  geom_bar(mapping = aes(x = support, 
                         color = support, 
                         fill = support)) +
  facet_grid(. ~ region)
```
By getting more granular, we are able to determine that Trump is more favored in the Midwest and the South. In the Northeast, Trump did not get any majority support in any of the states, and while the West did have some states where he had the majority, Clinton was overall the most favored


### Part 4 - Inference


### Part 5 - Conclusion


### References
Majumder, Maimuna. “Higher Rates Of Hate Crimes Are Tied To Income Inequality.” FiveThirtyEight, FiveThirtyEight, 21 Apr. 2017, fivethirtyeight.com/features/higher-rates-of-hate-crimes-are-tied-to-income-inequality/.

### Appendix (optional)

Remove this section if you don't have an appendix