---
title: "Data606 - Week1Hw"
author: "Baron Curtin"
date: "February 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Env Setup
```{r envsetup, echo=FALSE}
library(tidyverse)
library(magrittr)
library(DATA606)
library(openintro)
```

### 1.8
* a.) Each row in the data matrix represents a different UK resident from the sample that was collected from the survey to study smoking habits.
* b.) 1,691 participants were included in the survey
* c.)
    + sex: Categorical (Non-Ordinal)
    + age: Numerical (Discrete)
    + marital: Categorial (Non-Ordinal)
    + grossIncome: Categorical (Non-Ordinal)
    + smoke: Categorical (Non-Ordinal)
    + amtWeekends: Numerical (Discrete)
    + amtWeekdays: Numerical (Discrete)
    
### 1.10
* a.) The population of interest is children between the ages of 5 and 15. The sample in this study is the 160 children chosen between the ages of 5 and 15.
* b.) I do not believe that this study can be generalized to the population as there are too many variables in the sample. Age is the biggest variable as there is no way to guarantee honesty amongst either group of repsondents. In order to determine a causal relationship, the experiment would need to be redesigned taking a sample of each age one at a time, and conduct the test that way. The sample could also skew the results as there could be different distributions amongst the age groups in the sample.

### 1.28
* a.) I do not believe that we can determine causality from this study. There may be a correlation between smoking and dementia later in life, however there may be other factors involved that are actually causing the dementia.
* b.) The conclusion from the study are not reliable and can contain bias. Teachers and parents will have bias towards certain children. There may be another variable confounding the results of the survey and the results are therefore unreliable.

### 1.36
* a.) This kind of study is a prospective study as mental health is examined as time goes on or "events unfold".
* b.) The treatment group is the group that exercises. The control group is the group that does not exercise.
* c.) The study makes use of blocking and the blocking variable is age.
* d.) The excerpt does not make clear that each group does not know what the other is doing, however if they do not, blinding is used. This would only be a single-blind study.
* e.) The results may not be used to determine causality because there are other confounding variables that are not taken into account such as genetics and general health. If any conclusions can be made, they would have to be limited to the strata as opposed to the population at large.
* f.) I would have reservations about the study proposal because there is not enough blocking used to establish causality.

### 1.48
```{r 1.48}
vals <- data.frame(c(57,66,69,71,72,73,74,77,78,78,79,79,81,81,82,83,83,88,89,94))

ggplot(vals, aes(x = "", y = vals)) + geom_boxplot()
```

### 1.50
* a.) The histogram is approximately normally distributed and symmetrical, unimodal, and matches the boxplot labelled (2).
* b.) The histogram is symmetrical, multimodal, and matches the boxplot labelled (3).
* c.) The histogram is skewed right, and matches the boxplot labelled (1).

### 1.56
* a.) The distribution is right skewed, the median would best represent a typical observation, and IQR would best represent the variability of the data.
* b.) The distribution is symmetrical, the mean would best represent a typical observation, and standard deviation would best represent the variability of the data.
* c.) The distribution is right skewed, the median would best represent a typical observation, and IQR would best represent the variability of the data.
* d.) The distribution is right skewed, the median would best represent a typical observation, and IQR would best represent the variability of the data.

### 1.70
```{r 1.70}
library(openintro)
data(heartTr)

data <- heartTr
```
* a.) Based on the mosaic plot, there is an increase in survival if the patient got a transplant. While it is not completely dependent upon the transplant, there is some level of dependency based on the increase in survivors that were in the treatment group.
* b.) Based on the box plot, the heart transplant was more effective in generating a longer lifespan than those that did not receive the treatment.
* c.)
```{r 1.70c}
treatment <- data %>%
  filter(transplant == "treatment")
tProp <- (sum(treatment$survived == "dead") / nrow(treatment)) %>%
  print

control <- data %>%
  filter(transplant == "control")
tCont <- (sum(control$survived == "dead") / nrow(control)) %>%
  print
```
* d.)
    * i.) The claim being tested is does the heart transplant treatment increase lifespan?
    * ii.) We write alive on 28 cards representing patients who were alive at the end of the study, and dead on 79 cards representing patients who were not. Then, we shuffle these cards and split them into two groups: one group of size 69 representing treatment, and another group of size 34 representing control. We calculate the difference between the proportion of dead cards in the treatment and control groups (treatment -
control) and record this value. We repeat this 100 times to build a distribution centered at . Lastly, we calculate the fraction of simulations where the simulated differences in proportions are the proportion of patients in treatment that died minus the proportion of patients in control that died. If this fraction is low, we conclude that it is unlikely to have observed such an outcome by chance and that the null hypothesis should be rejected in favor of the alternative.
  * iii.) The simulation results show that the difference in proportion is likely not due by chance and this the transplant is effective.


    
